State of the Practice for Lattice Boltzmann Method Software

### 2022-10-02 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/ResubmissionResponseToReviewer.tex

\documentclass[12pt]{casletter}

\oddsidemargin -10mm
\evensidemargin -10mm
\textwidth 180mm
\textheight 200mm
%\renewcommand\baselinestretch{0.85}

\usepackage{xcolor}

\pagestyle {plain}
\pagenumbering {arabic}

\email{smiths@mcmaster.ca}
\telephone{(905) 525-9140 ext.\ 24131}
\www{{http://}www.cas.mcmaster.ca/\~\@smiths}

\signature{Spencer Smith}

\address{Dr.\ Spencer Smith, Associate Professor\\
Computing and Software Department\\
Faculty of Engineering\\
McMaster University\\
1280 Main Street West\\
Hamilton, Ontario, Canada L8S 4K1}%  \\ Phone: 525-9140 Ext.~27929 \\E-mail:
                                  %  smiths@mcmaster.ca}

\begin{document}

\begin{letter}{~\\
    ~\\
    ~\\
    ~\\
    Dr.\ Eugenio O\~{n}ate\\
    Editor-in-Chief\\
    Archives of Computational Methods in Engineering\\
    ~\\
    {\bf Re: Revisions to ARCO-D-22-00308R1, State of the Practice for Lattice
    Boltzmann Method Software}}

  \opening {Dear Dr.\ O\~{n}ate:}

  Thank you and the reviewer for the feedback on our submission.  In response to
  your e-mail, dated August 26, 2022, we have revised the paper to incorporate
  the requested revisions.  We provide a summary below.  We have also included
  with the revised submission a version of the paper showing all additions and
  deletions.

  We would like to thank you and the reviewer for their thoughtful and
  constructive comments.

  \textbf{Reviewer \#1}

  \begin{enumerate}
  \item The anonymization of the authors poses some challenges in evaluating
  the paper because the citations for the methodology and the raw data
  underlying the study have been anonymized away. \medskip

    \emph{We have removed the anonymization in the resubmission.  \smallskip}

  \item I think it would be useful to have slightly more by way of explanation
  here -- to provide a bit more understanding of the process followed for
  readers who choose not to pursue the details provided in the reference.  Two
  key steps to put a little more flesh around would be (1) how are the entries
  of the measurement template (Fig 2) turned into a single numerical score for
  the section, and (2) how is the aggregate AHP score obtained?  As to the first
  point, for example, I look at the Installability Overall Impression scores,
  which I initially took to be a numerical score for installability, but in the
  AHP Installability scores shown in Fig 3, there is no correlation between the
  number in the measurement template and the final score. \medskip

    \emph{Response. \smallskip}

  \item [I]n Fig 2, I see responses for "Descriptive error messages?" of yes, no,
  and n/a.  I have a hard time imagining what n/a could mean in this context.
  Presumably the installer either provides descriptive error messages or it does
  not?  Is this explained in one of the anonymized references or in the raw
  dataset, which I can't access either due to anonymization. \medskip

    \emph{Response. \smallskip}

  \item AHP is left completely to outside references, and I think it would be
  beneficial to have at least a brief description here. \medskip

    \emph{Response. \smallskip}

  \item I think there was a statement that AHP weights all of the qualities
  equally, but in the graphs of the scores for various qualities (e.g., Figs
  3-6), there are dramatic differences in the ranges of the scores given to the
  packages -- anywhere from 0.06 as the top score to nearly 0.16.  The highest
  overall score is ~0.075 (Fig 12). So what is the *theoretical* top score in
  any category and overall?  Are there some categories in which even the
  highest-scoring packages are quite low on the absolute scale?  If so, I would
  think such situations deserve some discussion. \medskip

    \emph{Response. \smallskip}

  \end{enumerate}

  \closing{Best regards,~\newline} \vspace{-29mm}
  \includegraphics[scale=0.75]{Signature.pdf}

\end {letter}

\end{document}

### 2022-10-03 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
updated paper to reflect updated figures

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 108 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of 20 questions). After measuring, we rank the software using the Analytic
    Hierarchy Process (AHP). Our ranking is roughly consistent with GitHub stars
    ranking, suggesting at least a partial correlation between the use of best
    practices and popularity. We find the state of the practice to be healthy,
    with 67\% of the measured packages ranking in the top five for at least one
    quality, the majority of LBM generated artifacts corresponding to general
    recommendations from research software developers, common use of version
    control (67\% of packages) and the adoption of a quasi-agile development
    process.  Areas of best practice to potentially improve include adoption of
    continuous integration, API documentation and enforcement of programming
    style guides.  We interviewed four developers to gain insight into their
    current pain points. Identified challenges include lack of development time,
    lack of funding, and difficulty with ensuring correctness.  Developers are
    addressing these pain points by designing for change, circumventing the
    oracle problem and prioritizing documentation and usability. For future
    improvements we suggest the following: employing linters, conducting
    rigorous peer reviews, writing and submitting more papers on software,
    growing the number of contributors by following current recommendations for
    open source projects, and augmenting the theory manuals to include more
    requirements specification relevant information.

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages. Software packages with a higher score (ESPResSo, MechSys, OpenLB,
DL\_MESO, Palabos) tend to have one set of linear installation instructions,
written as if the person doing the installation has none of the dependencies
installed. The instructions often list compatible operating system versions and
include instructions for installing dependencies. The top-ranked packages often
incorporate some automation of the installation process. The number of
dependencies a package has does not correlate with a higher score. The ability
to validate the installation process, say through tutorials or test cases, leads
to a higher score. Apparently active work on a project improves installability
since the top seven ranked packages are all classified as alive.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the ranking for surface
correctness and verifiability. Software packages with a higher score tend to
have theory documentation. They also explicitly use at least one document
generation tool that helps with verification, thus building confidence in
correctness. The top ranked software packages all include getting started
tutorials, and most of these include expected output. The two top ranked
packages, Ludwig and Musubi, indicate the use of unit testing in their
documentation. They also incorporated CI in the development process.
Furthermore, eight of the top 10 ranked packages are alive.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. All the top
five ranked packages have tutorials. The package pyLBM broke during tutorial
testing, but a descriptive error message helped in recovery. Nine of the top 10
packages are alive.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All the software packages that installed
correctly passed this test. They did not crash and usually output descriptive
error messages. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
their runtime robustness. All successfully installed software packages that
require a plain text input file correctly handled an unexpected change to these
input files, including a replacement of new lines with carriage returns.
Following the trend of the other qualities, nine of the top 10 ranked packages
are alive.

@@

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50\% of their issues closed. All the top five ranked packages have most of
their issues closed. Alive packages (11 use issue tracking) have 64\% of their
issues closed, while dead packages (3 use issue tracking) have 71\% of their
issues closed. Table~\ref{gitrepodata} presents this information. With respect
to the host for the git repository, 13 packages use GitHub and two (Palabos,
waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS
for issue tracking and version control, and seven do not appear to use any issue
tracking system.  Assuming the use of GitHub, GitLab or CVS implies the use of
version control, 16 of 24 (67\%) of projects use version control.

@@

Table~\ref{gitrepodata} shows the percentage of code that is comments, in
decreasing order. We assume maintainability improves for packages with a higher
percentage of comments. Comments represent more than 10\% of code files in 16
packages, and the average percentage of code comments is about 14\%. All the top
five overall ranked packages have more than the average. LUMA has only 0.2\%
comments, the fewest of any package. However, this package also has the most
lines of source code, with over four million. The next largest package is \esp{}
with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to manage their issue tracking, having most of their issues
closed. Maintainable packages by our ranking have more than 10\% of the code as
comments. Four of the top five ranked packages are alive.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages. Software packages with a high score have thousands of source code
files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modelling. For this reason, a comparison with other software packages is
not on a level field. Nine of the top 10 ranked packages are alive.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also have
detailed and easy to access software release notes. Four of the top five ranked
packages are alive.

@@

Notable exceptions to this trend occur for three projects that rank high by star
count, but relatively low in the ``follow best practices'' ranking: TCLB (star
rank 3, our rank 12), lettuce (star rank 4, our rank 14) and \esp{} (star rank
5, our rank 15).  Part of the discrepancy is because our process ranks all
projects, but the star measure does not apply to all projects.  Also, our
process equally weights all qualities (Section~\ref{Sec_OverallQuality}), but
for real users some qualities will be more important than others. Discrepancies
between our measures and the star measures may also be due to inaccuracy with
using stars to approximate popularity, because of how people use stars, because
young projects have less time to accumulate stars, and because stars represent
the community's feeling in the past more than current preferences
\citep{Szulik2017}. Older packages (like \esp{} and TCLB) have had more time to
accumulate stars and watches. Young projects have less time to accumulate stars,
like HemeLB, which only has 12 stars (at the time of measurement) having
recently moved from a different GitHub repository.  Lettuce is a young project
(released in 2019) with a relatively high number of stars. The discrepancy in
rankings may be due to a mismatch between the time we manually measured lettuce
and when we counted the stars. The late addition of Musubi means a gap between
when we measured projects and the time when our automated data (like stars) was
collected. A final reason for inconsistencies between our ranking and the
community's ranking is that, as for consumer products, more factors influence
popularity than just quality.

\begin{table}[ht!]
	\begin{center}
		\begin{tabular}{ p{3cm}p{1.25cm}p{1.75cm}p{1.75cm}p{1.75cm}p{1.75cm} }
			\toprule
			Name & Our Ranking & Repository Stars & Repository Star Rank &
			Repository Watches & Repository Watch Rank\\
			\midrule
			ESPResSo & 1 & 145 & 2 & 19& 2\\
			Ludwig & 2 & 27 & 8 & 6 & 7\\
			Palabos & 3 & 34 & 6 & GitLab & GitLab\\
			OpenLB & 4 & N/A & N/A & N/A& N/A\\
			LUMA & 5 & 33 & 7 & 12 & 4\\
			pyLBM & 6 & 95 & 3 & 10 & 5\\
			DL\_MESO (LBE) & 7 & N/A & N/A & N/A & N/A\\
			Musubi & 8 & N/A & N/A & N/A & N/A\\
			Sailfish & 9 & 186 & 1 & 41& 1\\
			waLBerla & 10 & 20 & 9 & GitLab & GitLab\\
			laboetie & 11 & 4 & 13 & 5 & 8\\
			TCLB & 12 & 95 & 3 & 16 & 3\\
			MechSys & 13 & N/A & N/A & N/A & N/A\\
			lettuce & 14 & 48 & 4 & 5 & 8\\
			\esp & 15 & 35 & 5 & 12 & 4\\
			MP-LABS & 16 & 12 & 11 & 2 & 9\\			
			SunlightLB & 17 & N/A & N/A & N/A & N/A\\
			LB3D & 18 & N/A & N/A & N/A & N/A\\			
			LIMBES & 19 & N/A & N/A & N/A & N/A\\
			LB2D-Prime & 20 & N/A & N/A & N/A & N/A\\		
			HemeLB & 21 & 12 & 11 & 12 & 4\\
			lbmpy & 22 &  11 & 12 & 2 & 9\\	
			LB3D-Prime & 23 & N/A & N/A & N/A & N/A\\	
			LatBo.jl & 24 & 17 & 10 & 8 & 6\\			
			\bottomrule
		\end{tabular}
		\caption{Repository Ranking Metrics (Sorted by Our Ranking)} \label{repometrics}
	\end{center}
\end{table}

@@

The top four packages (ESPREsSo, Ludwig, Palabos, and OpenLB) include most of
the common and uncommon artifacts, but not many of the rare artifacts.  They
collectively have all the common artifacts, except Palabos does not have theory
notes, and OpenLB does not appear to use version control.  The top four packages
have most of the uncommon artifacts. Only one them (Ludwig) is missing design
information, and only one (Palabos) does not have a user manual.  However, for
Palabos there was a broken link on the package website indicating that such an
artifact might exist. The developers later fixed this broken link, but this is
not reflected in our data because it was not present at the time of data
collection (mid-2020). Despite the broken link, Palabos does have a detailed and
informative website. The top four ranked packages do not have many of the rare
artifacts. None of them have any explicit API documentation. Three of them
(ESPREsSo, Ludwig, Palabos) have information on contributing to the project and
two (OpenLB, Palabos) have an FAQ section or forum. One (OpenLB) has
verification and validation notes, and a video guide for the software.

@@

Current LBM software mostly follows best practices, as illustrated by
Table~\ref{topperformerstable}, which summarizes the top five performers for
each quality and for the overall quality. Taking the union of all entries in the
table, we have 16 top performers: ESPREsSo, MechSys, OpenLB, DL\_MESO, Sailfish,
Ludwig, Musubi, pyLBM, waLBerla, laboetie, lettuce, \esp, LUMA, MP-LABS,
Palabos, and HemeLB.  Therefore, two thirds (67\%) of the 24 packages
appear in the top five for at least one quality.  Almost all entries are alive,
except for three packages: laboetie, MP-LABS, and HemeLB.  Alive packages are in
a healthy state with respect to adopting best practices.

\begin{table}[ht!]
	\begin{center}
		\begin{tabular}{ p{3cm}p{1.9cm}p{1.9cm}p{1.9cm}p{1.9cm}p{1.9cm} }
			\toprule
			Quality & Ranked 1st & Ranked 2nd & Ranked 3rd & Ranked 4th & Ranked
			5th\\
			\midrule
			Installability & ESPResSo & MechSys & OpenLB & DL\_MESO (LBE) &
			Sailfish\\
			\addlinespace[0.4cm]
			\pbox{3.0cm}{Surface Correctness \\ and Verifiability} & Ludwig &
			Musubi & pyLBM & OpenLB & waLBerla\\
			\addlinespace[0.4cm]
			Surface Reliability & DL\_MESO (LBE) & laboetie* & lettuce & MechSys
			& Musubi \\
			\addlinespace[0.4cm]
			Surface Robustness & DL\_MESO (LBE) & \esp & lettuce & LUMA &
			OpenLB \\
			\addlinespace[0.4cm]
			Surface Usability & laboetie* & ESPResSo & OpenLB & Ludwig & Sailfish\\
			\addlinespace[0.4cm]
			Maintainability & Ludwig & MP-LABS* & \esp & ESPResSo & Palabos\\
			\addlinespace[0.4cm]
			Reusability & ESPResSo & waLBerla & \esp & Sailfish & HemeLB*\\
			\addlinespace[0.4cm]
			\pbox{3.0cm}{Surface \\Understandability} & LUMA & Palabos & pyLBM &
			DL\_MESO (LBE)& MechSys\\
			\addlinespace[0.4cm]
			\pbox{3.0cm}{Visibility and \\Transparency} & Ludwig & Palabos &
			LUMA & ESPResSo & MP-LABS*\\
			\addlinespace[0.4cm]
			Overall Quality & ESPResSo & Ludwig & Palabos & OpenLB & LUMA\\
			\bottomrule
		\end{tabular}
		\caption{Top performers for each quality (all packages were alive at the
		time of measurement, except for those marked by an asterisk (*)) (Sorted
		by Order of Quality Measurement)}
		\label{topperformerstable}
	\end{center}
\end{table}

>>>>

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 108 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of 20 questions). After measuring, we rank the software using the Analytic
    Hierarchy Process (AHP). Our ranking is roughly consistent with GitHub stars
    ranking, suggesting at least a partial correlation between the use of best
    practices and popularity. We find the state of the practice to be healthy,
    with 87.5\% of the measured packages ranking in the top five for at least one
    quality, the majority of LBM generated artifacts corresponding to general
    recommendations from research software developers, common use of version
    control (67\% of packages) and the adoption of a quasi-agile development
    process.  Areas of best practice to potentially improve include adoption of
    continuous integration, API documentation and enforcement of programming
    style guides.  We interviewed four developers to gain insight into their
    current pain points. Identified challenges include lack of development time,
    lack of funding, and difficulty with ensuring correctness.  Developers are
    addressing these pain points by designing for change, circumventing the
    oracle problem and prioritizing documentation and usability. For future
    improvements we suggest the following: employing linters, conducting
    rigorous peer reviews, writing and submitting more papers on software,
    growing the number of contributors by following current recommendations for
    open source projects, and augmenting the theory manuals to include more
    requirements specification relevant information.

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages. Software packages with a higher score (ESPResSo, MechSys, OpenLB,
DL\_MESO, Palabos) tend to have one set of linear installation instructions,
written as if the person doing the installation has none of the dependencies
installed. The instructions often list compatible operating system versions and
include instructions for installing dependencies. The top-ranked packages often
incorporate some automation of the installation process. The number of
dependencies a package has does not correlate with a higher score. The ability
to validate the installation process, say through tutorials or test cases, leads
to a higher score. Apparently active work on a project improves installability
since the top six ranked packages are all classified as alive.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the ranking for surface
correctness and verifiability. Software packages with a higher score tend to
have theory documentation. They also explicitly use at least one document
generation tool that helps with verification, thus building confidence in
correctness. The top ranked software packages all include getting started
tutorials, and most of these include expected output. The two top ranked
packages, Ludwig and Musubi, indicate the use of unit testing in their
documentation. They also incorporated CI in the development process.
Furthermore, all of the top 5 ranked packages are alive.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. The package pyLBM broke during tutorial testing, but a descriptive error message helped in recovery. Ten of the top 11 packages are alive.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All the software packages that installed
correctly passed this test. They did not crash and usually output descriptive
error messages. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
their runtime robustness. All successfully installed software packages that
require a plain text input file correctly handled an unexpected change to these
input files, including a replacement of new lines with carriage returns.

@@

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50\% of their issues closed. All the top four ranked packages have most of
their issues closed. Alive packages (11 use issue tracking) have 64\% of their
issues closed, while dead packages (3 use issue tracking) have 71\% of their
issues closed. Table~\ref{gitrepodata} presents this information. With respect
to the host for the git repository, 13 packages use GitHub and two (Palabos,
waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS
for issue tracking and version control, and seven do not appear to use any issue
tracking system.  Assuming the use of GitHub, GitLab or CVS implies the use of
version control, 16 of 24 (67\%) of projects use version control.

@@

Table~\ref{gitrepodata} shows the percentage of code that is comments, in
decreasing order. We assume maintainability improves for packages with a higher
percentage of comments. Comments represent more than 10\% of code files in 16
packages, and the average percentage of code comments is about 14\%. All the top
four overall ranked packages have more than the average. LUMA has only 0.2\%
comments, the fewest of any package. However, this package also has the most
lines of source code, with over four million. The next largest package is \esp{}
with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to manage their issue tracking, having most of their issues
closed. Maintainable packages by our ranking have more than 10\% of the code as
comments. All of the top four ranked packages are alive.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages. Software packages with a high score have thousands of source code
files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modelling. For this reason, a comparison with other software packages is
not on a level field. Seven of the top 8 ranked packages are alive.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also have
detailed and easy to access software release notes. Five of the top six ranked
packages are alive.

@@

Notable exceptions to this trend occur for three projects that rank high by star
count, but relatively low in the ``follow best practices'' ranking: TCLB (star
rank 3, our rank 10), lettuce (star rank 4, our rank 12) and \esp{} (star rank
5, our rank 15).  Part of the discrepancy is because our process ranks all
projects, but the star measure does not apply to all projects.  Also, our
process equally weights all qualities (Section~\ref{Sec_OverallQuality}), but
for real users some qualities will be more important than others. Discrepancies
between our measures and the star measures may also be due to inaccuracy with
using stars to approximate popularity, because of how people use stars, because
young projects have less time to accumulate stars, and because stars represent
the community's feeling in the past more than current preferences
\citep{Szulik2017}. Older packages (like \esp{} and TCLB) have had more time to
accumulate stars and watches. Young projects have less time to accumulate stars,
like HemeLB, which only has 12 stars (at the time of measurement) having
recently moved from a different GitHub repository.  Lettuce is a young project
(released in 2019) with a relatively high number of stars. The discrepancy in
rankings may be due to a mismatch between the time we manually measured lettuce
and when we counted the stars. The late addition of Musubi means a gap between
when we measured projects and the time when our automated data (like stars) was
collected. A final reason for inconsistencies between our ranking and the
community's ranking is that, as for consumer products, more factors influence
popularity than just quality.

\begin{table}[ht!]
	\begin{center}
		\begin{tabular}{ p{3cm}p{1.25cm}p{1.75cm}p{1.75cm}p{1.75cm}p{1.75cm} }
			\toprule
			Name & Our Ranking & Repository Stars & Repository Star Rank &
			Repository Watches & Repository Watch Rank\\
			\midrule
			ESPResSo & 1 & 145 & 2 & 19& 2\\
			Ludwig & 2 & 27 & 8 & 6 & 7\\
			Palabos & 3 & 34 & 6 & GitLab & GitLab\\
			LUMA & 4 & 33 & 7 & 12 & 4\\
			OpenLB & 5 & N/A & N/A & N/A& N/A\\
			pyLBM & 6 & 95 & 3 & 10 & 5\\
			Musubi & 7 & N/A & N/A & N/A & N/A\\
			waLBerla & 8 & 20 & 9 & GitLab & GitLab\\
			DL\_MESO (LBE) & 9 & N/A & N/A & N/A & N/A\\
			Sailfish & 9 & 186 & 1 & 41& 1\\
			TCLB & 10 & 95 & 3 & 16 & 3\\
			laboetie & 11 & 4 & 13 & 5 & 8\\
			lettuce & 12 & 48 & 4 & 5 & 8\\
			MechSys & 13 & N/A & N/A & N/A & N/A\\
			\esp & 14 & 35 & 5 & 12 & 4\\
			LB2D-Prime & 15 & N/A & N/A & N/A & N/A\\
			LB3D & 16 & N/A & N/A & N/A & N/A\\
			SunlightLB & 16 & N/A & N/A & N/A & N/A\\
			HemeLB & 17 & 12 & 11 & 12 & 4\\
			MP-LABS & 17 & 12 & 11 & 2 & 9\\
			LIMBES & 18 & N/A & N/A & N/A & N/A\\
			lbmpy & 19 &  11 & 12 & 2 & 9\\
			LB3D-Prime & 20 & N/A & N/A & N/A & N/A\\
			LatBo.jl & 21 & 17 & 10 & 8 & 6\\
			\bottomrule
		\end{tabular}
		\caption{Repository Ranking Metrics (Sorted by Our Ranking)} \label{repometrics}
	\end{center}
\end{table}

@@

The top three packages (ESPREsSo, Ludwig, Palabos) include most of
the common and uncommon artifacts, but not many of the rare artifacts.  They
collectively have all the common artifacts, except Palabos does not have theory
notes.  The top three packages
have most of the uncommon artifacts. Only one them (Ludwig) is missing design
information, and only one (Palabos) does not have a user manual.  However, for
Palabos there was a broken link on the package website indicating that such an
artifact might exist. The developers later fixed this broken link, but this is
not reflected in our data because it was not present at the time of data
collection (mid-2020). Despite the broken link, Palabos does have a detailed and informative website. The top three ranked packages do not have many of the rare artifacts. None of them have any explicit API documentation. All of them have information on contributing to the project and
one (Palabos) has an FAQ section or forum.

@@

Current LBM software mostly follows best practices, as illustrated by
Table~\ref{topperformerstable}, which summarizes the top five performers for
each quality and for the overall quality. Taking the union of all entries in the
table, we have 21 top performers: ESPREsSo, MechSys, OpenLB, DL\_MESO, Palabos, Sailfish, Ludwig, Musubi, pyLBM, waLBerla, laboetie, lettuce, LUMA, TCLB, \esp, LB2D-Prime, LB3D-Prime, LIMBES, SunlightLB, HemeLB, MP-LABS.  Therefore, seven eights (87.5\%) of the 24 packages
appear in the top five for at least one quality.  Most entries are alive.  Alive packages are in
a healthy state with respect to adopting best practices.

\begin{table}[ht!]
	\begin{center}
		\begin{tabular}{ p{3cm}p{3cm}p{2.1cm}p{2.1cm}p{2.1cm}p{2.1cm} }
			\toprule
			Quality & Ranked 1st & Ranked 2nd & Ranked 3rd & Ranked 4th & Ranked
			5th\\
			\midrule
			Installability & ESPResSo, MechSys, OpenLB & & & DL\_MESO (LBE),
			Palabos, Sailfish &\\
			\addlinespace[0.4cm]
			\pbox{3.0cm}{Surface Correctness \\ and Verifiability} & Ludwig &
			Musubi, pyLBM & & ESPResSo, waLBerla &\\
			\addlinespace[0.4cm]
			Surface Reliability & Musubi, DL\_MESO (LBE), ESPResSo, laboetie*, lettuce, LUMA, MechSys, OpenLB, Palabos, pyLBM, TCLB, & & & & \\
			\addlinespace[0.4cm]
			Surface Robustness & DL\_MESO, ESPResSo, \esp, laboetie*, LB2D-Prime*, LB3D-Prime*, lettuce, LIMBES*, Ludwig, LUMA, MechSys, OpenLB, Palabos, pyLBM, Sailfish, SunlightLB*, TCLB, waLBerla & & & & \\
			\addlinespace[0.4cm]
			Surface Usability & ESPResSo, laboetie*, Ludwig, OpenLB, Sailfish & & & & \\
			\addlinespace[0.4cm]
			Maintainability & ESPResSo, Ludwig, Palabos, waLBerla & & & & \esp \\
			\addlinespace[0.4cm]
			Reusability & ESPResSo, waLBerla & & \esp, HemeLB*, Musubi, OpenLB, Palabos, Sailfish & &\\
			\addlinespace[0.4cm]
			\pbox{3.0cm}{Surface \\Understandability} & LUMA & DL\_MESO (LBE), lettuce, Musubi, Palabos & & & \\
			\addlinespace[0.4cm]
			\pbox{3.0cm}{Visibility and \\Transparency} & Ludwig & ESPResSo, LUMA, Palabos & & & MP-LABS*\\
			\addlinespace[0.4cm]
			Overall Quality & ESPResSo & Ludwig & Palabos & LUMA & OpenLB \\
			\bottomrule
		\end{tabular}
		\caption{Top performers for each quality (all packages were alive at the
		time of measurement, except for those marked by an asterisk (*)) (Sorted
		by Order of Quality Measurement)}
		\label{topperformerstable}
	\end{center}
\end{table}
